Labor Markets and Technological Change: Evidence from Electronic Health Records

### 2021-10-20 Paper/Paper.tex
Updates from Overleaf

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,hyperref,array,xcolor,multicol,verbatim,mathpazo}
\usepackage[normalem]{ulem}
\usepackage[pdftex]{graphicx}
\usepackage{fullpage}

@@

\title{Physician Response to EHR Implementation in Hospitals}
\author{Hanna Glenn}
%\date{\today}

@@

Using CMS Shared Patient Data from 2009-2015, I construct a panel of physician-hospital pairs that captures a specific working relationship between the pair, most likely the physician working in the hospital. I link the hospitals to AHA survey data for information on EHR use and then aggregate the data to the physician level. To measure the working decisions of physicians, I utilize (1) physician level information, also from the shared patient data, that captures total working life (referrals to labs, pharmacies, other physicians, etc), and (2) an office-based physician level dataset to observe whether physicians are moving from hospitals to private practice. For my analysis, I use a difference-in-difference estimation where treatment is defined as physicians' exposure to an EHR at any of the hospitals they work in. A key identifying assumption in this analysis is that hospitals choose to implement EHRs not based on any factors that are also correlated with physician working decisions.

\newpage
\printbibliography
\end{document}

>>>>

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,hyperref,array,xcolor,multicol,verbatim,mathpazo}
\usepackage[normalem]{ulem}
\usepackage[pdftex]{graphicx}
\usepackage{fullpage}
\usepackage{import}
\usepackage{adjustbox}

@@

\title{Labor Markets and Technological Change: Evidence from Electronic Health Records}
\author{Hanna Glenn}
%\date{\today}

@@

Using CMS Shared Patient Data from 2009-2015, I construct a panel of physician-hospital pairs that captures a specific working relationship between the pair, most likely the physician working in the hospital. I link the hospitals to AHA survey data for information on EHR use and then aggregate the data to the physician level. This physician-level data gives information on an individuals' level of exposure to electronic health records. To measure the working decisions of physicians, I utilize (1) physician level information, also from the shared patient data, that captures total working life (referrals to labs, pharmacies, other physicians, etc), and (2) an office-based physician level data set to observe whether physicians are moving from hospitals to private practice. For my analysis, I use a difference-in-difference estimation where treatment is physicians' exposure to an EHR at any of the hospitals they work in. A key identifying assumption in this analysis is that hospitals choose to implement EHRs not based on any factors that are also correlated with physician working decisions.

Paragraph about findings

This paper contributes to two main portions of literature: the effects of health information technology and the factors of physician labor market decisions. The effects of EHR implementation on costs and quality of healthcare prior to 2010 have been studied extensively. Despite many case studies and hospital-level analysis that indicate large positive effects on health (\cite{Buntin2011TheResults}), economic research has showed that the technology has only improved health outcomes for patients with severe conditions, but have not led to improvements on average (\cite{Agha2014TheCare}; \cite{McCullough2016HealthCoordination}; \cite{Meyerhoefer}). Further, there have been no significant cost decreases due to EHRs in the short or long run (CITE COST STUDIES). However, the long run is not fully realized, so there may still be potential benefits realized in the future.

The relationship between this technology and physicians, which are the primary users of the technology, is ...

\section{Background}

\section{Data}

\section{Empirical Analysis}

\section{Results}

\newpage

\import{Objects/}{continuous.tex}

\newpage

\import{Objects/}{indicator.tex}

\newpage

\import{Objects/}{continuous_old.tex}

\newpage

\import{Objects/}{indicator_old.tex}

\newpage

\end{document}

### 2021-10-20 Presentations/ThirdYearPaperpres.tex
Updates from Overleaf

\title{Hospital Information Technology and Physician Response}

@@

\begin{frame}{What is an EHR?}

\begin{itemize}
    \item Computer-program technology that hospitals purchase
    \item Performance of EHR depends on level purchased:
    \begin{itemize}
        \item All include electronic patient records
        \item More complex systems allow for decision making assistance
        \item More recently, patient interfaces are being developed
    \end{itemize}
\end{itemize}

\begin{frame}[fragile]{HIT had Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item A 2005 estimate finds possible cost reduction of hundreds of billions of dollars (\cite{hillestad2005})
\end{itemize}
\end{alertblock}

\begin{alertblock}{Quality Improvement}
\begin{itemize}
    \item Improved efficiency, patient safety improvements, physicians have decision support that could prevent unnecessary complications, etc.
    \item Significant policy push for EHR implementation with these goal in mind: HITECH Act, 2008 provided financial incentive for hospitals to implement EHRs \nocite{hitech}
\end{itemize}
\end{alertblock}

@@

\begin{frame}[fragile]{What does this mean for physicians?}
\begin{itemize}
    \item Initially, this means physicians spend more time entering information into a computer and learning a new software
    \item Loss of autonomy as EHRs have progressed
    \item Physicians have experienced significant burn-out from EHRs
\end{itemize}

@@

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.4]{graphics/News Clip1.PNG}
\end{center}
\end{frame}

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.5]{graphics/News Clip2.PNG}
\end{center}
\end{frame}

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.45]{graphics/News Clip3.PNG}
\end{center}
\end{frame}

@@

Does EHR implementation in hospitals affect labor market decisions of physicians, specifically where, and how much they work?

>>>>

\title{Labor Markets and Technological Change: Evidence from Electronic Health Records}

@@

\begin{frame}{What is an EHR?}
\begin{itemize}
    \item Computer-program technology that hospitals purchase
    \vspace{3mm}
    \item What EHR is used for depends on level purchased:
    \vspace{2mm}
    \begin{itemize}
        \item All include electronic patient records
        \vspace{2mm}
        \item More complex systems allow for decision making assistance
        \vspace{2mm}
        \item More recently, patient interfaces are being developed
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{HIT: Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item Possible cost reduction of hundreds of billions of dollars \\ (\cite{hillestad2005})
\end{itemize}
\end{alertblock}

\begin{alertblock}{Quality Improvement}
\begin{itemize}
    \item Improved efficiency, patient safety improvements, physicians have decision support that could prevent unnecessary complications, etc.
    \item Significant policy push for EHR implementation: HITECH Act, 2008 provided financial incentive for hospitals to implement EHRs \nocite{hitech}
\end{itemize}
\end{alertblock}

@@

\begin{frame}[fragile]{What does this mean for physicians?}
In the short run:
\begin{itemize}
    \item Physicians spend more time entering information into a computer and learning a new software
    \item Physicians have experienced significant burn-out from using and learning EHRs
\end{itemize}

In the long run:
\begin{itemize}
    \item Could make daily tasks easier or harder
    \item Continued loss of autonomy as technology progresses
\end{itemize}

@@

\begin{frame}[noframenumbering]{Physician Burnout in the Media}
\begin{center}
    \includegraphics[scale=.4]{graphics/News Clip1.PNG}
\end{center}
\end{frame}

\begin{frame}[noframenumbering]{Physician Burnout in the Media}
\begin{center}
    \includegraphics[scale=.5]{graphics/News Clip2.PNG}
\end{center}
\end{frame}

\begin{frame}[noframenumbering]{Physician Burnout in the Media}
\begin{center}
    \includegraphics[scale=.45]{graphics/News Clip3.PNG}
\end{center}
\end{frame}

@@

Did EHR implementation in hospitals affect labor market decisions of physicians, specifically where, and how much they work?
